---
title: "Targeted CAMK2D Dataset Discovery for Cardiac Research"
author: "CAMK2D Research Consortium"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
    df_print: paged
  pdf_document:
    toc: true
params:
  focus_area: "both"  # "aFIB", "HF", or "both"
  target_species: "all"  # "human", "mouse", or "all" 
  min_samples: 10  # Lowered from 20 to include more studies
  require_controls: FALSE  # Allow observational studies
  validate_camk2d: TRUE
  download_expression: TRUE
  output_dir: "output"
  max_datasets: 30  # Moderate increase to balance performance
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

# Load required libraries
suppressPackageStartupMessages({
  library(tidyverse)
  library(plotly)
  library(openxlsx)
  library(DT)
  library(knitr)
  library(kableExtra)
})

# Load our enhanced search functions
source("functions/targeted_camk2d_search.R")
source("functions/expression_data_validator.R")

# Create output directory
if (!dir.exists(params$output_dir)) {
  dir.create(params$output_dir, recursive = TRUE)
}
```

# Executive Summary

This document implements **targeted CAMK2D dataset discovery** specifically aligned with our research proposal focusing on:

- **Primary Objective**: CAMK2D role in atrial fibrillation (aFIB) and heart failure (HF)
- **Data Requirements**: Actual expression matrices suitable for differential gene expression analysis
- **Quality Validation**: CAMK2D expression levels, case-control design, adequate sample sizes
- **Bioinformatics Ready**: Expression data, phenotype annotations, and analysis summaries

**Key Enhancement**: This system moves beyond generic dataset discovery to **proposal-specific, validated datasets** with actual expression data ready for immediate bioinformatics analysis.

# 1. Research Focus Configuration

## 1.1 Analysis Parameters

```{r parameters}
cat("🎯 CAMK2D Research Configuration\n")
cat("================================\n")
cat("Focus Area:", params$focus_area, "\n")
cat("Target Species:", params$target_species, "\n") 
cat("Minimum Samples:", params$min_samples, "\n")
cat("Require Controls:", params$require_controls, "\n")
cat("Validate CAMK2D:", params$validate_camk2d, "\n")
cat("Download Expression Data:", params$download_expression, "\n")
cat("Maximum Datasets:", params$max_datasets, "\n\n")

# Display research objectives alignment
objectives_table <- data.frame(
  Objective = c(
    "CAMK2D role in aFIB",
    "CAMK2D role in HF", 
    "Expression analysis ready",
    "Case-control validation",
    "CAMK2D expression validation",
    "Bioinformatics pipeline ready"
  ),
  Status = c(
    ifelse(params$focus_area %in% c("aFIB", "both"), "✅ Enabled", "⚠️ Disabled"),
    ifelse(params$focus_area %in% c("HF", "both"), "✅ Enabled", "⚠️ Disabled"),
    ifelse(params$download_expression, "✅ Enabled", "❌ Disabled"),
    ifelse(params$require_controls, "✅ Required", "⚠️ Optional"),
    ifelse(params$validate_camk2d, "✅ Required", "⚠️ Optional"),
    ifelse(params$download_expression, "✅ Ready", "❌ Not Ready")
  )
)

kable(objectives_table, caption = "Research Objectives Alignment") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# 2. Targeted Dataset Discovery

## 2.1 Proposal-Specific Search Execution

```{r targeted-discovery}
cat("🚀 Executing targeted CAMK2D dataset discovery...\n\n")

# Execute targeted search
discovery_results <- discover_targeted_camk2d_datasets(
  focus_area = params$focus_area,
  species = params$target_species,
  min_samples = params$min_samples,
  require_controls = params$require_controls
)

# Display results
discovered_datasets <- discovery_results$datasets
search_summary <- discovery_results$summary

if (nrow(discovered_datasets) > 0) {
  
  cat("📊 Discovery Results Summary:\n")
  cat("   Total relevant datasets:", search_summary$total_datasets, "\n")
  cat("   High CAMK2D relevance:", search_summary$high_relevance, "\n")
  cat("   Medium CAMK2D relevance:", search_summary$medium_relevance, "\n")
  cat("   Suitable for DGE analysis:", search_summary$dge_suitable, "\n")
  cat("   Primary analysis candidates:", search_summary$primary_candidates, "\n")
  cat("   Average sample size:", search_summary$avg_sample_size, "\n\n")
  
  # Display top datasets
  top_datasets <- discovered_datasets %>%
    head(min(params$max_datasets, nrow(discovered_datasets))) %>%
    dplyr::select(GEO_Accession, Title, Organism, Sample_Count, CAMK2D_Relevance, 
           Priority_Class, Recommendation, Total_Relevance_Score)
  
  kable(top_datasets, 
        caption = "Top CAMK2D-Relevant Datasets Discovered",
        col.names = c("Accession", "Title", "Species", "Samples", 
                     "CAMK2D Relevance", "Priority", "Recommendation", "Score")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), font_size = 10) %>%
    scroll_box(width = "100%")
  
} else {
  cat("❌ No relevant datasets discovered with current criteria\n")
  cat("💡 Suggestions:\n")
  cat("   - Broaden species criteria (set target_species = 'all')\n") 
  cat("   - Reduce minimum samples (set min_samples = 10)\n")
  cat("   - Disable control requirement (set require_controls = FALSE)\n")
}
```

## 2.2 Search Strategy Analysis

```{r search-strategy-analysis}
if (!is.null(discovery_results$search_strategies) && nrow(discovered_datasets) > 0) {
  
  # Analyze search strategy effectiveness
  strategy_effectiveness <- discovered_datasets %>%
    count(Search_Strategy, name = "Datasets_Found") %>%
    arrange(desc(Datasets_Found))
  
  cat("🔍 Search Strategy Effectiveness:\n")
  for (i in 1:nrow(strategy_effectiveness)) {
    strategy <- strategy_effectiveness$Search_Strategy[i]
    count <- strategy_effectiveness$Datasets_Found[i]
    description <- discovery_results$search_strategies[[strategy]]$description
    cat("   ", strategy, ":", count, "datasets -", description, "\n")
  }
  cat("\n")
  
  # Visualize strategy effectiveness
  strategy_plot <- ggplot(strategy_effectiveness, 
                         aes(x = reorder(Search_Strategy, Datasets_Found), 
                             y = Datasets_Found)) +
    geom_col(fill = "#3498db", alpha = 0.8) +
    coord_flip() +
    labs(title = "Search Strategy Effectiveness",
         subtitle = "Number of relevant datasets found per strategy",
         x = "Search Strategy", 
         y = "Datasets Found") +
    theme_minimal()
  
  print(strategy_plot)
}
```

# 3. Expression Data Validation

## 3.1 Download and Validate Expression Data

```{r expression-validation}
validated_datasets <- list()
validation_summary <- list()

if (params$download_expression && nrow(discovered_datasets) > 0) {
  
  cat("📥 Downloading and validating expression data...\n\n")
  
  # Select top datasets for validation (limit to prevent timeout)
  top_datasets_for_validation <- discovered_datasets %>%
    filter(Priority_Class %in% c("Highest", "High")) %>%
    arrange(desc(Total_Relevance_Score)) %>%
    head(5)  # Limit to top 5 to prevent timeout
  
  if (nrow(top_datasets_for_validation) == 0) {
    top_datasets_for_validation <- head(discovered_datasets, 3)
  }
  
  cat("🎯 Validating", nrow(top_datasets_for_validation), "high-priority datasets\n\n")
  
  for (i in 1:nrow(top_datasets_for_validation)) {
    dataset <- top_datasets_for_validation[i, ]
    geo_accession <- dataset$GEO_Accession
    
    cat("Processing", i, "of", nrow(top_datasets_for_validation), ":", geo_accession, "\n")
    
    # Download and validate
    validation_result <- download_and_validate_expression(
      geo_accession = geo_accession,
      output_dir = file.path(params$output_dir, "expression_data"),
      validate_camk2d = params$validate_camk2d,
      min_expression = 1
    )
    
    if (!is.null(validation_result) && validation_result$suitable) {
      validated_datasets[[geo_accession]] <- validation_result
      cat("✅", geo_accession, "validated and ready for analysis\n")
    } else {
      reason <- if (!is.null(validation_result)) validation_result$reason else "Download failed"
      cat("❌", geo_accession, "failed validation:", reason, "\n")
    }
    
    cat("\n")
  }
  
  # Summarize validation results
  n_validated <- length(validated_datasets)
  n_attempted <- nrow(top_datasets_for_validation)
  
  cat("📊 Expression Data Validation Summary:\n")
  cat("   Datasets attempted:", n_attempted, "\n")
  cat("   Successfully validated:", n_validated, "\n")
  cat("   Validation success rate:", round(n_validated/n_attempted*100, 1), "%\n\n")
  
  if (n_validated > 0) {
    # Create validation summary table
    validation_summary_data <- map_dfr(validated_datasets, function(dataset) {
      data.frame(
        GEO_Accession = dataset$geo_accession,
        Genes = dataset$analysis_summary$n_genes,
        Samples = dataset$analysis_summary$n_samples,
        Groups = dataset$analysis_summary$n_groups,
        CAMK2D_Expression = round(dataset$analysis_summary$camk2d_mean_expr, 3),
        Suitable_for_DGE = dataset$analysis_summary$suitable_for_dge,
        Missing_Data_Percent = round(dataset$analysis_summary$missing_data_percent, 1),
        stringsAsFactors = FALSE
      )
    })
    
    kable(validation_summary_data,
          caption = "Expression Data Validation Results",
          col.names = c("Accession", "Genes", "Samples", "Groups", 
                       "CAMK2D Expr", "DGE Ready", "Missing %")) %>%
      kable_styling(bootstrap_options = c("striped", "hover"))
  }
  
} else {
  cat("⚠️ Expression data download disabled or no datasets to validate\n")
}
```

## 3.2 CAMK2D Expression Analysis

```{r camk2d-expression-analysis}
if (length(validated_datasets) > 0) {
  
  cat("🧬 CAMK2D Expression Analysis Across Validated Datasets\n\n")
  
  # Extract CAMK2D expression data
  camk2d_expression_summary <- map_dfr(validated_datasets, function(dataset) {
    camk2d_val <- dataset$camk2d_validation
    
    if (!is.null(camk2d_val) && camk2d_val$camk2d_detected) {
      result <- data.frame(
        Dataset = dataset$geo_accession,
        CAMK2D_Mean_Expression = round(camk2d_val$camk2d_mean_expression, 3),
        CAMK2D_Max_Expression = round(camk2d_val$camk2d_max_expression, 3),
        Features_Found = length(camk2d_val$camk2d_features_found),
        stringsAsFactors = FALSE
      )
      
      # Add differential expression info if available
      if (!is.null(camk2d_val$differential_expression) && 
          camk2d_val$differential_expression$testable) {
        diff_expr <- camk2d_val$differential_expression
        result$Group1 <- diff_expr$group1_name
        result$Group2 <- diff_expr$group2_name
        result$Group1_Mean <- round(diff_expr$group1_mean, 3)
        result$Group2_Mean <- round(diff_expr$group2_mean, 3)
        result$Log2_Fold_Change <- round(diff_expr$log2_fold_change, 3)
        result$P_Value <- format(diff_expr$p_value, scientific = TRUE, digits = 3)
        result$Significant <- diff_expr$significant
      }
      
      return(result)
    }
    return(NULL)
  })
  
  if (nrow(camk2d_expression_summary) > 0) {
    kable(camk2d_expression_summary,
          caption = "CAMK2D Expression Analysis Summary",
          col.names = c("Dataset", "Mean Expr", "Max Expr", "Features", 
                       "Group 1", "Group 2", "Group1 Mean", "Group2 Mean", 
                       "Log2 FC", "P Value", "Significant")) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), font_size = 9) %>%
      scroll_box(width = "100%")
    
    # Visualize CAMK2D expression levels
    if (nrow(camk2d_expression_summary) >= 2) {
      expr_plot <- ggplot(camk2d_expression_summary, 
                         aes(x = reorder(Dataset, CAMK2D_Mean_Expression), 
                             y = CAMK2D_Mean_Expression)) +
        geom_col(fill = "#e74c3c", alpha = 0.8) +
        coord_flip() +
        labs(title = "CAMK2D Expression Levels Across Datasets",
             subtitle = "Mean expression levels in validated datasets",
             x = "Dataset", 
             y = "CAMK2D Mean Expression") +
        theme_minimal()
      
      print(expr_plot)
    }
  }
}
```

# 4. Dataset Quality and Recommendation

## 4.1 Comprehensive Dataset Scoring

```{r final-dataset-scoring}
if (nrow(discovered_datasets) > 0) {
  
  cat("📊 Final Dataset Quality Assessment and Recommendations\n\n")
  
  # Enhanced scoring incorporating validation results
  final_scored_datasets <- discovered_datasets %>%
    mutate(
      Expression_Validated = GEO_Accession %in% names(validated_datasets),
      Expression_Score = ifelse(Expression_Validated, 10, 0),
      Final_Score = Total_Relevance_Score + Expression_Score,
      Final_Recommendation = case_when(
        Expression_Validated & Priority_Class == "Highest" ~ "PRIMARY - Ready for Analysis",
        Expression_Validated & Priority_Class == "High" ~ "VALIDATION - Ready for Analysis", 
        Expression_Validated ~ "SUPPLEMENTARY - Analysis Ready",
        Priority_Class == "Highest" ~ "HIGH PRIORITY - Needs Validation",
        Priority_Class == "High" ~ "MEDIUM PRIORITY - Needs Validation",
        TRUE ~ "LOW PRIORITY - Consider if needed"
      )
    ) %>%
    arrange(desc(Final_Score))
  
  # Display final recommendations
  recommendation_summary <- final_scored_datasets %>%
    dplyr::select(GEO_Accession, Title, Sample_Count, CAMK2D_Relevance, 
           Expression_Validated, Final_Score, Final_Recommendation) %>%
    head(10)
  
  kable(recommendation_summary,
        caption = "Final Dataset Recommendations for CAMK2D Research",
        col.names = c("Accession", "Title", "Samples", "CAMK2D Relevance",
                     "Validated", "Final Score", "Recommendation")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), font_size = 9) %>%
    scroll_box(width = "100%")
  
  # Recommendation distribution
  rec_distribution <- table(final_scored_datasets$Final_Recommendation)
  
  cat("💡 Recommendation Summary:\n")
  for (rec in names(rec_distribution)) {
    cat("   ", rec, ":", rec_distribution[rec], "datasets\n")
  }
  cat("\n")
}
```

## 4.2 Research Priority Matrix

```{r priority-matrix}
if (nrow(discovered_datasets) > 0) {
  
  # Create priority matrix for research planning
  priority_matrix <- final_scored_datasets %>%
    mutate(
      Research_Phase = case_when(
        str_detect(Final_Recommendation, "PRIMARY") ~ "Phase 1: Primary Analysis",
        str_detect(Final_Recommendation, "VALIDATION") ~ "Phase 2: Validation",
        str_detect(Final_Recommendation, "SUPPLEMENTARY") ~ "Phase 3: Meta-Analysis",
        TRUE ~ "Phase 4: Future Consideration"
      ),
      Data_Readiness = ifelse(Expression_Validated, "Analysis Ready", "Needs Processing"),
      Proposal_Alignment = case_when(
        CAMK2D_Relevance == "High" & str_detect(Search_Strategy, "afib|hf") ~ "Perfect Match",
        CAMK2D_Relevance == "High" ~ "High Alignment", 
        CAMK2D_Relevance == "Medium" ~ "Moderate Alignment",
        TRUE ~ "Low Alignment"
      )
    ) %>%
    count(Research_Phase, Data_Readiness, Proposal_Alignment) %>%
    pivot_wider(names_from = c(Data_Readiness, Proposal_Alignment), 
                values_from = n, 
                values_fill = 0)
  
  kable(priority_matrix,
        caption = "Research Priority Matrix") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```

# 5. Bioinformatics Pipeline Readiness

## 5.1 Analysis-Ready Datasets Summary

```{r analysis-ready-summary}
analysis_ready_datasets <- names(validated_datasets)

if (length(analysis_ready_datasets) > 0) {
  
  cat("🧬 Bioinformatics Pipeline Ready Datasets\n")
  cat("=========================================\n\n")
  
  for (accession in analysis_ready_datasets) {
    dataset <- validated_datasets[[accession]]
    
    cat("📊 Dataset:", accession, "\n")
    cat("   📁 Expression file:", basename(dataset$saved_files$expression_file), "\n")
    cat("   🏷️  Phenotype file:", basename(dataset$saved_files$phenotype_file), "\n")
    cat("   📝 Feature file:", basename(dataset$saved_files$feature_file), "\n")
    cat("   🧬 Genes:", dataset$analysis_summary$n_genes, "\n")
    cat("   👥 Samples:", dataset$analysis_summary$n_samples, "\n")
    cat("   📊 Groups:", dataset$analysis_summary$n_groups, "\n")
    cat("   ✅ CAMK2D expressed:", dataset$analysis_summary$camk2d_expressed, "\n")
    cat("   🔬 DGE ready:", dataset$analysis_summary$suitable_for_dge, "\n")
    cat("\n")
  }
  
  # Generate analysis commands
  cat("💻 Suggested Analysis Commands:\n")
  cat("# Load datasets for analysis\n")
  for (accession in analysis_ready_datasets) {
    cat("# Dataset:", accession, "\n")
    cat(paste0("expr_", gsub("[^A-Za-z0-9]", "_", accession), " <- read_csv('", 
               file.path(params$output_dir, "expression_data", paste0(accession, "_expression.csv")), "')\n"))
    cat(paste0("pheno_", gsub("[^A-Za-z0-9]", "_", accession), " <- read_csv('", 
               file.path(params$output_dir, "expression_data", paste0(accession, "_phenotypes.csv")), "')\n"))
    cat("\n")
  }
  
} else {
  cat("⚠️  No datasets are currently ready for bioinformatics analysis\n")
  cat("💡 To get analysis-ready datasets:\n")
  cat("   1. Set download_expression = TRUE\n")
  cat("   2. Ensure adequate datasets meet validation criteria\n")
  cat("   3. Check CAMK2D expression requirements\n")
}
```

# 6. Export and Documentation

## 6.1 Comprehensive Results Export

```{r export-results}
if (nrow(discovered_datasets) > 0) {
  
  cat("📊 Exporting comprehensive results...\n")
  
  # Create Excel workbook
  wb <- createWorkbook()
  
  # Sheet 1: Discovery Results
  addWorksheet(wb, "Discovery_Results")
  writeData(wb, "Discovery_Results", final_scored_datasets)
  
  # Sheet 2: Validation Summary
  if (exists("validation_summary_data") && nrow(validation_summary_data) > 0) {
    addWorksheet(wb, "Validation_Results")
    writeData(wb, "Validation_Results", validation_summary_data)
  }
  
  # Sheet 3: CAMK2D Expression
  if (exists("camk2d_expression_summary") && nrow(camk2d_expression_summary) > 0) {
    addWorksheet(wb, "CAMK2D_Expression")
    writeData(wb, "CAMK2D_Expression", camk2d_expression_summary)
  }
  
  # Sheet 4: Analysis Summary
  analysis_overview <- data.frame(
    Metric = c("Total Datasets Discovered", "High CAMK2D Relevance", 
               "Expression Data Validated", "Analysis Ready Datasets",
               "Primary Candidates", "Validation Candidates"),
    Value = c(nrow(discovered_datasets), 
              sum(discovered_datasets$CAMK2D_Relevance == "High"),
              length(validated_datasets),
              sum(final_scored_datasets$Expression_Validated),
              sum(str_detect(final_scored_datasets$Final_Recommendation, "PRIMARY")),
              sum(str_detect(final_scored_datasets$Final_Recommendation, "VALIDATION")))
  )
  
  addWorksheet(wb, "Analysis_Overview")
  writeData(wb, "Analysis_Overview", analysis_overview)
  
  # Save Excel file
  excel_file <- file.path(params$output_dir, 
                         paste0("CAMK2D_Targeted_Discovery_", format(Sys.Date(), "%Y%m%d"), ".xlsx"))
  saveWorkbook(wb, excel_file, overwrite = TRUE)
  
  cat("✅ Results exported to:", excel_file, "\n")
  
  # Also save key datasets as CSV
  csv_file <- file.path(params$output_dir, 
                       paste0("CAMK2D_Discovery_Results_", format(Sys.Date(), "%Y%m%d"), ".csv"))
  write_csv(final_scored_datasets, csv_file)
  
  cat("✅ Dataset details saved to:", csv_file, "\n")
}
```

# 7. Research Recommendations

## 7.1 Next Steps for CAMK2D Analysis

```{r research-recommendations}
cat("🔬 CAMK2D Research Recommendations\n")
cat("==================================\n\n")

if (length(analysis_ready_datasets) > 0) {
  cat("✅ IMMEDIATE ACTION ITEMS:\n")
  cat("1. Begin differential expression analysis with", length(analysis_ready_datasets), "validated datasets\n")
  cat("2. Focus on CAMK2D expression patterns across disease groups\n")
  cat("3. Identify CAMK2D co-expressed genes for pathway analysis\n") 
  cat("4. Cross-validate findings across multiple datasets\n\n")
  
  cat("📊 SUGGESTED ANALYSIS WORKFLOW:\n")
  cat("1. Quality control and normalization of expression data\n")
  cat("2. Differential expression analysis (case vs control)\n")
  cat("3. CAMK2D-centered gene network analysis\n")
  cat("4. Pathway enrichment focusing on cardiac function\n")
  cat("5. Cross-species validation if mouse data available\n\n")
}

if (nrow(discovered_datasets) > length(analysis_ready_datasets)) {
  unvalidated_count <- nrow(discovered_datasets) - length(analysis_ready_datasets)
  cat("⚠️  FUTURE VALIDATION NEEDED:\n")
  cat("•", unvalidated_count, "additional relevant datasets identified\n")
  cat("• Consider validating high-priority unprocessed datasets\n")
  cat("• Focus on datasets with 'HIGH PRIORITY - Needs Validation' status\n\n")
}

# Proposal alignment check
cat("🎯 PROPOSAL ALIGNMENT CHECK:\n")
alignment_check <- data.frame(
  Requirement = c(
    "aFIB-focused datasets",
    "HF-focused datasets", 
    "CAMK2D expression validated",
    "Case-control design",
    "Adequate sample sizes",
    "Expression data ready"
  ),
  Status = c(
    ifelse(any(str_detect(discovered_datasets$Search_Strategy, "afib")), "✅ Found", "❌ Missing"),
    ifelse(any(str_detect(discovered_datasets$Search_Strategy, "hf")), "✅ Found", "❌ Missing"),
    ifelse(length(validated_datasets) > 0, "✅ Validated", "⚠️  Partial"),
    ifelse(params$require_controls, "✅ Required", "⚠️  Optional"), 
    ifelse(any(discovered_datasets$Sample_Count >= params$min_samples), "✅ Met", "❌ Below threshold"),
    ifelse(length(analysis_ready_datasets) > 0, "✅ Ready", "❌ Not ready")
  )
)

kable(alignment_check, caption = "Proposal Requirements Status") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

# Final Summary

```{r final-summary}
cat("📋 CAMK2D Targeted Discovery Summary\n")
cat("=====================================\n")
cat("Analysis completed on:", format(Sys.Date(), "%Y-%m-%d"), "\n")
cat("Focus area:", params$focus_area, "\n")
cat("Target species:", params$target_species, "\n\n")

cat("📊 Key Results:\n")
cat("• Datasets discovered:", ifelse(exists("discovered_datasets"), nrow(discovered_datasets), 0), "\n")
cat("• Expression data validated:", length(validated_datasets), "\n")
cat("• Analysis-ready datasets:", length(analysis_ready_datasets), "\n")
cat("• High CAMK2D relevance:", ifelse(exists("discovered_datasets"), sum(discovered_datasets$CAMK2D_Relevance == "High"), 0), "\n\n")

if (length(analysis_ready_datasets) > 0) {
  cat("🚀 READY FOR ANALYSIS:\n")
  cat("Your CAMK2D research can proceed immediately with validated expression data\n")
  cat("Next step: Begin differential gene expression analysis\n")
} else {
  cat("⚠️  ACTION REQUIRED:\n")
  cat("Enable expression data download to get analysis-ready datasets\n")
  cat("Set download_expression = TRUE and re-run analysis\n")
}
```

---

*Targeted CAMK2D Dataset Discovery Report generated on `r Sys.Date()` with proposal-specific validation and expression data readiness.*